---
title: "Lab 02 - Plastic waste"
author: "Colin Li"
date: "1/17/2023"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

###warm up

```{r}
View(plastic_waste)
#Source Editor, Console, Global Environment, and Plots
#240 observations
#NA means missing data
```

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.
Did you expect this result? You might consider doing some research on Trinidad and Tobago to see why plastic waste per capita is so high there, or whether this is a data error.

What can you say about how the continents compare to each other in terms of their plastic waste per capita?

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```

### Exercise 2

#Describe why we defined the color and fill of the curves by mapping aesthetics of the plot but we defined the alpha level as a characteristic of the plotting geom.
###It looks like color and fill just automatically choose colors for different continents, and alpha controls the density of the graph


```{r plastic-waste-density}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.2)

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()

```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
  
### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements



