---
title: "Lab 02 - Plastic waste"
author: "Colin Li"
date: "1/17/2023"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

###warm up

```{r}
View(plastic_waste)
#Source Editor, Console, Global Environment, and Plots
#240 observations
#NA means missing data
```

### Exercise 1

#Trinidad and Tobago import tons of plastic every year, so it makes sense and it's not data error. 

#It looks like many African and Asian countries have less plastic waste per capita, one or maybe a few North American countries have a lot plastic waste per capita

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap, fill = continent)) +
  geom_histogram(binwidth = 0.2)

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent)) +
  geom_density()

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```

### Exercise 2


###It looks like color and fill just automatically choose colors for different continents. Color will color the dots/lines (if there are any). Fill just fills the colors in the graph. and alpha controls the density of the graph


```{r plastic-waste-density}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, fill = continent)) +
  geom_density(alpha = 0.2)

ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()

```

### Exercise 3

What do the violin plots reveal that box plots do not?
#Violin plots also show density and data distribution (e.g., normality)

What features are apparent in the box plots but not in the violin plots?
#the medians. 

```{r plastic-waste-violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

### Exercise 4


```{r}


plastic_waste %>%
  summarize(cor(plastic_waste_per_cap, total_pop))



```



### Exercise 5


```{r plastic-waste-mismanaged}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap)) +
  geom_point()

```

### Exercise 6

#it looks like a weak positive relationship and it looks like Oceania has less plastic waste per capita but they mismanage plastic waste more. Europe did the best in terms of mismanage plastic waste. 

```{r plastic-waste-mismanaged-continent}



ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = mismanaged_plastic_waste_per_cap,
                     color = continent)) +
  geom_point()

```

### Exercise 7

#They don't seem to be more strongly linearly associated compare to the previous graph
```{r plastic-waste-population-total}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = total_pop,
                     color = continent)) +
  geom_point()

```

```{r plastic-waste-population-coastal}

ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     y = coastal_pop,
                     color = continent)) +
  geom_point()


```

### Exercise 8

#I think it's identical except for the color
```{r recreate-viz}

library(ggsci)
plastic_waste_f <- plastic_waste %>% 
  filter(plastic_waste_per_cap < 3)

plastic_waste_f$Coastal_pop_pro <- plastic_waste_f$coastal_pop/plastic_waste_f$total_pop

ggplot(data = plastic_waste_f, 
       mapping = aes(x = Coastal_pop_pro, 
                     y = plastic_waste_per_cap)) +
  ggtitle("Plastic waste vs. coastal population proportion", "by continent") + 
  xlab("Coastal population proportion (Coastal / total population)") + 
  ylab("Plastic waste per capita") + 
  geom_point(data = plastic_waste_f, aes(color=continent)) +
  geom_smooth(data = plastic_waste_f, color="black")+ theme(
  panel.background = element_rect(fill = NA),
  panel.grid.major = element_line(colour = "grey"),
) + scale_color_aaas()


```


